---
title: "Mixing cover crops suppresses weeds and roto-till reduces urban soil strength and improves infiltration"
author:
  - Naim Edwards ^a^*
  - Nicholas Medina ^b^
  - Elizabeth Asker ^a^
#   - name: Naim Edwards
#     email: edwar649@msu.edu
#     affiliation: Michigan State University
#     correspondence: yes
#   - name: Nicholas Medina
#     email: nmedina@umich.edu
#     affiliation: University of Michigan
#     correspondence: yes
#   - name: Elizabeth Asker
#     email: asker@msu.edu
#     affiliation: Michigan State University
# address:
#   - code: University of Michigan
#     address: Ecology and Evolutionary Biology, 1105 N University Ave, Ann Arbor, MI 48109
#   - code: Michigan State University
#     address: Agriculture and Natural Resources Extension, Detroit, MI
date: \scriptsize 
  ^a^ Agriculture and Natural Resources, Michigan State University, 16745 Lamphere St, Detroit, MI, USA 48219;  
  ^b^ Ecology and Evolutionary Biology, University of Michigan, 1105 N University Ave, Ann Arbor, MI, USA

# abstract: |

# keywords: "Keywords: urban, compact, weed, till, cover, water"
  
output: 
  # bookdown::word_document2:
  #   toc: FALSE
  #   global_numbering: TRUE #dev
  bookdown::pdf_document2:
    # number_sections: FALSE
    toc: FALSE
    keep_tex: TRUE
    global_numbering: TRUE #dev
    # template:
    #   ../../elsarticle-ecrc/ecrc-template.tex
    # ../../plot_latex_template/plos_latex_template.tex
    # includes:
    #   in-header: ../../elsarticle-ecrc/ecrc.sty
  # rticles::elsevier_article:
  #   keep_tex: TRUE
always_allow_html: TRUE
bibliography: 
  - ../../library.bib
  - ../../grateful-refs.bib
  - ../../software.bib
csl: ../../styles-master/elsevier-harvard.csl
# biblio-style: 
#   ../../elsarticle-ecrc/elsarticle-num.bst
  # plos2015.bst
# documentclass: elsarticle
# documentclass: ../../elsarticle-ecrc/elsarticle.cls
  
link-citations: TRUE
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage[left]{lineno}\linenumbers
  - \usepackage{setspace}\doublespace
# font-family: Calibri Light
---


<!-- Journals:  -->
<!-- - Elsevier--Soil & Tillage Research (IF=5, Elsevier),  -->
<!-- *- Land Degradation & Development (Research vs. Original, IF=5, Wiley), * -->
<!-- *- Urban Forestry & Urban Greening (1wk), or then [Wiley--]* -->
<!-- -- Urban Ecosystems (3.5, but 101d) -->
<!-- - ASA Agronomy Journal (?),  -->
<!-- - Ecology & Evolution (replicates), or lastly  -->
<!-- - PCI Ecology submission (45d) w/  -->
<!--   - pre-print archived in agriRxiv, ecoEvoRxiv, or bioRxiv. -->


```{r setupFull, echo=F}
options(digits = 1
        # knitr.table.format = "latex"
        );
# knitr::opts_chunk$set(fig.width = 3, fig.height = 3)
taxize::taxize_options(quiet = T, taxon_state_messages = T)
```


\
\

**Keywords**: urban agriculture; soil strength; weed suppression; roto-till; cover crop mix; soil infiltration  

\
\
  
***Corresponding author**: Naim Edwards, edwar649@msu.edu  
Author emails: edwar649@msu.edu, nmedina@umich.edu, askereli@msu.edu  
ORCIDs: Nicholas Medina, 0000-0001-5465-3988


\newpage


<!-- ### Highlights{-} -->

```{r intro, child='highlights.Rmd'}
```


\newpage


# Abstract{-}

```{r abs, child='abs.Rmd'}
```


\newpage


# Introduction

```{r intro, child='intro.Rmd'}
```


# Methods

```{r intro, child='meth.Rmd'}
```


# Results

```{r intro, child='res.Rmd'}
```


# Discussion

```{r intro, child='disc.Rmd'}
```


\newpage


# Funding{-}

*This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.*


# Declaration of interests{-}

Authors declare no conflicts of interest. 


# Acknowledgements{-}

Authors thank Brother Nature Produce, Earthworks Urban Farm, and Georgia Street Community Collective for early project input; site intern JH for field assistance; volunteer GV for site background help; and previous anonymous reviewers and peers KS, ZHF, and JK for discussion of initial drafts.


# Author contributions{-}

NE conceived, designed, and performed the study; NE and NM helped collect data; NM analyzed data; NE wrote the initial report. All authors wrote and revised second draft; NM wrote the third draft; all authors approved recent version.


# Data statement{-}

Code stored with Zenodo as [10.5281/zenodo.6800153](https://doi.org/10.5281/zenodo.6800153) *[@softwareMedinaEdwards22]* and linked to [github.com/nmedina17/must](github.com/nmedina17/must).


\newpage


# References
