---
title: "Combining medium-intensity tillage to improve soil compaction and infiltration with cover crop mixes to suppress weeds"
author:
  - name: Naim Edwards
    email: edwar649@msu.edu
    affiliation: Michigan State University
  - name: Nicholas Medina
    email: nmedina@umich.edu
    affiliation: University of Michigan
  - name: Elizabeth Asker
    email: asker@msu.edu
    affiliation: Michigan State University
address:
  - code: University of Michigan
    address: Ecology and Evolutionary Biology, Ann Arbor, MI
  - code: Michigan State University
    address: Agriculture and Natural Resources Extension, Detroit, MI

abstract: |
  Urban soils have been degraded by decades of industrial activities, but they also represent opportunities to improve food sovereignty for urban residents practicing urban agriculture. 
  Urban growers often rely on independent strategies of compost, tillage, or cover cropping, without distinguishing each's distinct benefits, ultimately informing an integrated approach. 
  This study examined how tillage methods representing various intensities and cover crop mixes targeting different functions affected agricultural variables including soil compaction, water infiltration rate, herbaceous weedy plant pressure, and crop yield. 
  Results showed that tractor-till significantly relieved compaction in deeper soils compared to roto-till, followed by roto-till compared to no-till.
  However, roto-till showed the fastest infiltration and tractor-till allowed the densest weeds, challenging ubiquitous use of tractor-till for urban agriculture.
  Results also showed that the mix of sorghum-sudangrass, buckwheat, and cowpea significantly reduced weed density and richness compared to all other mixes, and that the perennial plant mix significantly affected compaction, but without affecting soil water infiltration rates.
  Despite theses differences, radish yield was not yet affected by tillage.
  Overall, this study further informs urban soil agricultural management by analyzing specific benefits of common methods, advocating an integrated approach including both roto-till for compaction relief and water infiltration benefit, and mixed cover cropping for weed suppression.

keywords:
  - urban
  - compact
  - weed
  - till
  - cover
  - water
output: 
  # html_notebook
  bookdown::html_document2:
    number_sections: FALSE
always_allow_html: TRUE
bibliography: ../../library.bib
link-citations: TRUE
# csl: ../../peerj.csl
linenumbers: TRUE
---


Journals: 
- Elsevier--Soil & Tillage Research (IF=5, Elsevier), 
*- Land Degradation & Development (Research vs. Original, IF=5, Wiley), *
*- Urban Forestry & Urban Greening (1wk), or then [Wiley--]*
- ASA Agronomy Journal (?), 
- Ecology & Evolution (replicates), or lastly 
- PCI Ecology submission (45d) w/ 
  - pre-print archived in agriRxiv, ecoEvoRxiv, or bioRxiv.

<!-- # Abstract -->

<!-- ```{r abs, child='abs.Rmd'} -->
<!-- ``` -->


# Introduction

```{r intro, child='intro.Rmd'}
```


# Methods

```{r intro, child='meth.Rmd'}
```


# Results

```{r intro, child='res.Rmd'}
```


# Discussion

```{r intro, child='disc.Rmd'}
```


\newpage


# References
