---
title: Combining medium-intensity tillage to improve soil compaction and infiltration with cover crop mixes to suppress weeds
author:
  - name: Naim Edwards
    email: edwar649@msu.edu
    affiliation: Michigan State University
  - name: Nicholas Medina
    email: nmedina@umich.edu
    affiliation: University of Michigan
  - name: Elizabeth Asker
    email: asker@msu.edu
    affiliation: Michigan State University
address:
  - code: University of Michigan
    address: Ecology and Evolutionary Biology, 1105 N University Ave, Ann Arbor, MI 48109
  - code: Michigan State University
    address: Agriculture and Natural Resources Extension, Detroit, MI

abstract: |
  Urban soils have been degraded by decades of industrial activities, but they also represent opportunities to improve food sovereignty for urban residents practicing urban agriculture.
  Urban growers often use varying practices of compost, tillage, and cover cropping, but without distinguishing their benefits to optimize integrated approaches.
  This study examined how tillage methods representing various intensities and cover crop mixes targeting different functions affected agricultural variables including soil compaction, water infiltration rate, herbaceous weedy plant pressure, and crop yield. 
  Results showed that both roto- and tractor-till significantly affected compaction but not yield compared to no-till, and roto-till also improved infiltration, while tractor-till reached deeper soils but allowed denser weed growth.
  Mixing sorghum-sudangrass, buckwheat, and cowpea cover crops significantly reduced weed pressure compared to other mixtures, and perennials affected compaction but not soil water infiltration under no-till.
  These results reveal that medium-intensity tillage may offer more balanced trade-offs for initial management, and that cover crops can help reduce weeds under low-till strategies.
  Overall this study offers evidence detailing effects of various tillage and cover crop styles that can be of use for smallholder urban growers.

keywords:
  - urban
  - compact
  - weed
  - till
  - cover
  - water
output: 
  # html_notebook
  bookdown::pdf_book:
    # number_sections: FALSE
always_allow_html: FALSE
bibliography: 
- ../../library.bib
- ../../grateful-refs.bib
link-citations: TRUE
# csl: ../../peerj.csl
linenumbers: TRUE
fontsize: 12pt
geometry: margin=1in
---


<!-- Journals:  -->
<!-- - Elsevier--Soil & Tillage Research (IF=5, Elsevier),  -->
<!-- *- Land Degradation & Development (Research vs. Original, IF=5, Wiley), * -->
<!-- *- Urban Forestry & Urban Greening (1wk), or then [Wiley--]* -->
<!-- - ASA Agronomy Journal (?),  -->
<!-- - Ecology & Evolution (replicates), or lastly  -->
<!-- - PCI Ecology submission (45d) w/  -->
<!--   - pre-print archived in agriRxiv, ecoEvoRxiv, or bioRxiv. -->


`r options(digits = 1)`



<!-- # Abstract -->

<!-- ```{r abs, child='abs.Rmd'} -->
<!-- ``` -->


# Introduction

```{r intro, child='intro.Rmd'}
```


# Methods

```{r intro, child='meth.Rmd'}
```


# Results

```{r intro, child='res.Rmd'}
```


# Discussion

```{r intro, child='disc.Rmd'}
```


# Acknowledgements


Thanks to previous anonymous reviewers, and peer student (Kristel, Zach,) discussion of initial drafts.


# Funding


*This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.*


# Conflicts of interest


Authors declare no conflicts of interest. 


# Data statement


Data and code available at [nmedina17.github.com/must](nmedina17.github.com/must).


\newpage


# References
